---
title: "data_tidying"
author: "Sunny Siddique"
date: "September 25, 2018"
output: html_document
---

```{r}
library(tidyverse)
```

## Gather

PULSE data

```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
janitor::clean_names()
```

```{r}
pulse_tidy = pulse_data %>% 
  gather(key = "visit", value = "bdi_score", bdi_score_bl:bdi_score_12m)
```


Illusetrate 'separate':

```{r}
pulse_df = pulse_tidy %>% 
  separate(visit, into = c("bdi_str", "score_str", "visit"), sep = "_") %>% 
  select (-bdi_str, -score_str) %>% 
  mutate(visit = replace (visit, visit == "bl", "00m"))
```

```{r}
pulse_df
```

## Revisit FAS_litters

```{r}
litters_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day"), 3)
```


Learning Assessment: In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use  recode …). Is this version “tidy”?

```{r}
litters_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(litter_number, ends_with("weight")) %>% 
  gather(key = "gest_day", value = "weight", gd0_weight:gd18_weight) %>% 
  arrange(litter_number)
```

## Spread

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
```

Make it readable

```{r}
spread(analysis_result, key = time, value = mean) %>% 
  knitr::kable()

```














