---
title: "data_manipulation"
author: "Sunny Siddique"
date: "September 18, 2018"
output: html_document
---

```{r}
library(tidyverse)
```

## import fas csv file

Import my csv files to use as example

```{r}
litters_data = read_csv ("./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```

```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## Select variables

```{r}
select(litters_data, group, litter_number, gd0_weight)
select(litters_data, -litter_number)
```

## Example: selecting variables from the pups dataset

```{r}
select(pups_data, litter_number, sex, pd_ears)
```


## Filter observations

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)
```

## Filter to include only pups with sex 1 and filter to include only pups with pd walk less than 11 and sex 2.

```{r}
filter(pups_data, sex == 1)
filter(pups_data, pd_walk < 11, sex == 2)
```


## Mutate

Create variables
```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group)
       )
```


## Pipeline

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Use nested function calls

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

## piping

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight
  )
```

## Example practice of pipes
loads the pups data
cleans the variable names
filters the data to include only pups with sex 1
removes the PD ears variable
creates a variable that indicates whether PD pivot is 7 or more days

```{r}
read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  filter(sex == 1) %>% 
  select(-pd_ears) %>% 
  mutate(
    new_pd = pd_pivot >= 7)
```





